Proposal Defense

### 2025-11-25 presentation/proposal/proposal.qmd
Edit proposal

---
title: "Proposal Defense"
author:
  - name: Mona Mousavi$^*$
institute:
  - $^*$ University of Nebraska Lincoln
date: 06/12/2025
format:
  revealjs: default
  beamer:
    theme: Singapore
    navigation: horizontal
    section-titles: false
    slide-level: 2
    fontsize: "10pt"
header-includes: 
logo: nebraska-n.jpg
---

@@

# Background and Research Objectives

@@

+ On-farm precision experiments (OFPE) as a promising way (GPS technologies and variable-rate applicators).
  + Run on-farm agronomic trials and collect data
  + \textcolor{blue}{Estimate how yield responds to nitrogen based on observed characteristics}
  + Identify site-specific optimal nitrogen rates

@@

+ In all of the previous studies, the best model was selected based on their abilities to predict \textcolor{blue}{yield level} at the observation points (yield-based model selection).

@@

# Method

@@

Steps for a single iteration:

  + Step 1: Generate OFPE experiment data
  + Step 2: Identify the true best (most profitable VRA application rate) model among the five candidate models
  + Step 3: Select the best model among the 5 candidate models based on
    + yield (conventional approach)
    + local-EONR (our approach)
  + Step 4: Evaluate the model selection approaches

The five candidate models:
linear model, linear spatial error (SE) model, random forest (RF), boosted regression forest (BRF), and causal forest (CF).

## Data generation

@@

+ One possible scenario of the spatial variability in the field characteristics is shown.

@@

## The simulated OFPE field:
```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "95%"
knitr::include_graphics(
  here::here("presentation/proposal/g_field_map.png")
)
```

## True yield responses

@@

The true economically optimal N rate for each cell is then determined by:

$$
N_{j, i}^* \equiv \operatorname{argmax}_N\left[p \cdot f_{j, i}(N)-w \cdot N\right]
$$

+ $p$: the crop price, 
+ $w$ the nitrogen fertilizer price, 
+ $f_{j, i}(N)$: previously defined true yield response function for the specific cell.

## On-farm trial (nitrogen rate)

+ six N trial rates ($\{80, 130, 154, 184, 219, 270\}$ kg/ha.) followed a 6 × 6 Latin square design.

@@

## Yield data

@@

## independent variables

+ The true response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ and economically optimal N rates $N_{j, i}^*$ are not directly observable by producers or researchers.

@@

$$\beta = \beta^1 + \beta^2, \;\;
\beta^1 = \beta^{1,1} + \beta^{1,2}, \;\;\beta^2 = \beta^{2,1} + \beta^{2,2}$$ 
$$\tau = \tau^1 + \tau^2, \;\;
\tau^1 = \tau^{1,1} + \tau^{1,2}, \;\;\tau^2 = \tau^{2,1} + \tau^{2,2}$$ 
$$
\eta = \min(\eta^1, \eta^2), \;\;
\eta^1 = \min(\eta^{1,1}, \eta^{1,2}), \;\;
\eta^2 = \min(\eta^{2,1}, \eta^{2,2})
$$

# Identifying the true best model among the five candidate models

@@

## An example of the train-test split for each of the five folds

@@

## Chapter two:

### Causal methods for Site-specific yield response function estimation

# Background and Research Objectives

@@

+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response \textcolor{blue}{curve}.

@@

### Double-ML (binary treatment ($T$) case):

DML works fine in a \textcolor{blue}{binary} treatment ($T$) case (model assumes linearity in $T$):

@@

$$
\begin{aligned}
& Y_{res}= \\
& \begin{aligned}
\theta(X) \cdot T + g(X, W) + v 
- & \underbrace{E\{[\theta(x) \cdot T + g(X, W) + v] \mid X, W\}}= \\
& \textcolor{blue}{=} \textcolor{blue}{\theta(X) \cdot E(T \mid X, W) + g(X, W)} \\
& \textcolor{blue}{=}  \textcolor{blue}{\theta(X) \cdot h(X, W) + g(X, W)}
\end{aligned} \\
& \theta(X) \cdot T + g(X, W) + v 
- \theta(X) \cdot h(X, W) - g(X, W)= \\
& \theta(X) \cdot [\underbrace{T - h(X, W)}_{T_{res}}] + v \\
& \Rightarrow Y_{res} = \theta(X) \cdot T_{res} + v
\end{aligned}
$$

### Double-ML (continuous treatment ($T$) case):

DML fails in a \textcolor{blue}{continuous} treatment ($T$) case:

@@

$$
RMSE_{EONR}=\sqrt{\frac{1}{n} \sum_{i=1}^n\left(\widehat{EONR}_i-EONR_i\right)^2}
$$

>>>>

---
title: "Proposal Defense"
author:
  - name: Mona Mousavi
institute:
  - University of Nebraska Lincoln
date: 06/12/2025
format:
  revealjs:
    theme: [default, ./../custom.scss]
    fontsize: 1.4em
    callout-icon: false
    scrollable: true
    fig-dpi: 400
  beamer:
    theme: Singapore
    navigation: horizontal
    section-titles: false
    slide-level: 2
    fontsize: "10pt"
logo: nebraska-n.jpg
---

@@

@@

+ On-farm precision experiments (OFPE) as a promising way (GPS technologies and variable-rate applicators).
  + Run on-farm agronomic trials and collect data
  + <span style = "color: blue;">Estimate how yield responds to nitrogen based on observed characteristics</span>
  + Identify site-specific optimal nitrogen rates

@@

+ In all of the previous studies, the best model was selected based on their abilities to predict <span style = "color: blue;">yield level</span> at the observation points (yield-based model selection).

@@

## Method: Overview

@@

### Steps for a single iteration:

  + Step 1: Generate OFPE experiment data
  + Step 2: Identify the true best (most profitable VRA application rate) model among the five candidate models
  + Step 3: Select the best model among the 5 candidate models based on their abilities to predict
    + yield level (conventional approach)
    + local-EONR (our approach)
  + Step 4: Evaluate the model selection approaches

### The five candidate models:

+ Linear model 
+ Linear spatial error (SE) model 
+ Random forest (RF) 
+ Boosted regression forest (BRF) 
+ Causal forest (CF)

## Method: Data generation

@@

+ One possible scenario of the spatial variability in the field characteristics is shown below.

@@

```{r}
#| echo: false
#| fig-align: "center"
#| out-width: "95%"
knitr::include_graphics(
  here::here("presentation/proposal/g_field_map.png")
)
```

## Method: True yield responses

@@

## Method: On-farm trial (nitrogen rate)

Six N trial rates ($\{80, 130, 154, 184, 219, 270\}$ kg/ha.) following a 6 × 6 Latin square design.

@@

## Method: Yield data

@@

## Method: Independent variables

+ The true response parameters $\left(\alpha_{j, i}, \beta_{j, i}, \gamma_{j, i}, \tau_{j, i}, \eta_{j,i}\right)$ are not directly observable by producers or researchers.

@@

$$\beta = \beta^{1,1} + \beta^{1,2} + \beta^{2,1} + \beta^{2,2}$$ 
$$\tau = \tau^{1,1} + \tau^{1,2} + \tau^{2,1} + \tau^{2,2}$$ 
$$
\eta = \min(\min(\eta^{1,1}, \eta^{1,2}), \min(\eta^{2,1}, \eta^{2,2}))
$$

## Identifying the true best model among the five candidate models

@@

@@

# Chapter 2: Causal methods for Site-specific yield response function estimation

@@

+ Most causal machine learning methods (e.g., R-learner) are designed for binary treatments, whereas our objective is to recover the full yield response <span style = "color: blue;">curve</span>.

@@

## Method: Double-ML (binary treatment ($T$) case):

DML works fine in a <span style = "color: blue;">binary</span> treatment ($T$) case (model assumes linearity in $T$):

@@

$$
\begin{aligned}
& Y_{res} = \\
& \begin{aligned}
& {[\theta(X)\cdot T + g(X,W) + v] - [E\{ \theta(X)\cdot T + g(X,W) + v \mid X,W\}]= } \\
& {[\theta(X)\cdot T + g(X,W) + v] - [\theta(X)\cdot E(T\mid X,W) + g(X,W)]=} \\
&  {[\theta(X)\cdot T + g(X,W) + v] - [\theta(X)\cdot h(X,W) + g(X,W)]=} \\
&  \theta(X)\cdot [T - h(X,W)] + v =  \\
& \theta(X)\cdot T_{res} + v
\end{aligned}
\\[0.5em]
& \Rightarrow\quad Y_{res} = \theta(X)\cdot T_{res} + v
\end{aligned}
$$

## Method: Double-ML (continuous treatment ($T$) case):

DML fails in a <span style = "color: blue;">continuous</span> treatment ($T$) case:

@@

$$
RMSE_{EONR} = \sqrt{\frac{\sum_{i=1}^{n} (\widehat{EONR}_{i} - {EONR}_{i})^2}{n}}
$$
